Icicle: write once, run once

### 2016-05-25 2016fhpc/Main.tex
Mention modal type system in abstract
Wibbles to intro

\begin{abstract}
We present Icicle, a pure streaming query language which statically guarantees that multiple queries over the same input data stream will be fused. Icicle queries are compiled via a fold-based intermediate language to efficient C code. We present production benchmarks demonstrating significant speedup over existing queries written in R, as well as the widely used Unix tools \texttt{grep} and \texttt{wc}.

>>>>

\begin{abstract}
We present Icicle, a pure streaming query language which statically guarantees that multiple queries over the same input stream will be fused. We use a modal type system to ensure that fused queries can be computed in an incremental fashion, and a fold-based intermediate language to compile down to efficient C code. We present production benchmarks demonstrating significant speedup over existing queries written in R, as well as the widely used Unix tools \texttt{grep} and \texttt{wc}.

### 2016-05-25 2016fhpc/section/01-Introduction.tex
Wibbles to intro

A joint query like this can be converted to a back-end language such a SQL, but doing it by hand it tedious and error prone. As the three sub queries use different filter predicates we cannot use a single @SELECT@ statement and a @WHERE@ expression to implement the filter. We instead need to lift each predicate to an expression-level conditional:
\begin{code}
  SELECT COUNT(IF(open > close, 1, 0))
       , COUNT(IF(open < close, 1, 0))
       , SUM  (IF(open > close, open, 0))
       / COUNT(IF(open > close, 1, 0))
  FROM stocks;
\end{code}
As we see, the result of query fusion also tends to have many common sub expressions, and we wish to guarantee that the duplicates in the fused result are eliminated.

Joint queries such as the stocks example can be evaluated in a streaming, incremental fashion. As a counter example, suppose we have a table with two fields @key@ and @value@, and we wish to find the mean of values whose key matches the last one in the table. We might try something like:
\begin{code}
  table kvs { key : Int; value : Int }
  query avg = let k = last key
              in  filter (key == k) in mean value;
\end{code}
Unfortunately, although the \emph{result} we desire is computable, the \emph{algorithm} implied by the above query cannot be evaluated in a streaming fashion. While streaming through the table we always have access to the last key in the stream, but to perform a filter to select rows that match this key would require that we stream through the table again from the start. We need a better solution. With this in mind, the contributions of this paper are:
\begin{itemize}
\item
we present a type system that guarantees any two queries on the same input table can be fused, as well as allowing queries to be resumed as new data is received\sref{s:IcicleSource};

\item
we present an fold-based intermediate language, which allows the query fusion transformation to be a simple matter of appending two programs\sref{s:IcicleCore};

\item
and we present production benchmarks of our generated program, which compare favourably against standard unix utilities, and outperforms an existing R feature generation system by several orders of magnitude\sref{s:Benchmarks}.
\end{itemize}

>>>>

A joint query like this can be converted to a back-end language such as SQL, but doing it by hand it tedious and error prone. As the three sub queries use different filter predicates we cannot use a single @SELECT@ statement and a @WHERE@ expression to implement the filter. We instead need to lift each predicate to an expression-level conditional:
\begin{code}
  SELECT COUNT(IF(open > close, 1, 0))
       , COUNT(IF(open < close, 1, 0))
       , SUM  (IF(open > close, open, 0))
       / COUNT(IF(open > close, 1, 0))
  FROM stocks;
\end{code}
As we see, the result of query fusion also tends to have many common sub expressions, and we wish to guarantee that the duplicates in the fused result are eliminated.

Joint queries such as the stocks example can be evaluated in a streaming, incremental fashion. As a counter example, suppose we have a table with two fields @key@ and @value@, and we wish to find the mean of values whose key matches the last one in the table. We might try something like:
\begin{code}
  table kvs { key : Int; value : Int }
  query avg = let k = last key
              in  filter (key == k) in mean value;
\end{code}
Unfortunately, although the \emph{result} we desire is computable, the \emph{algorithm} implied by the above query cannot be evaluated in a streaming fashion. While streaming through the table we always have access to the last key in the stream, but to perform the filter to select rows that match this key would require that we stream through the table again from the start. We need a better solution. With this in mind, the contributions of this paper are:
\begin{itemize}
\item
We present a type system that guarantees any two queries on the same input table can be fused, as well as allowing queries to be resumed as new data is received\sref{s:IcicleSource};

\item
We present an fold-based intermediate language, which allows the query fusion transformation to be a simple matter of appending two programs\sref{s:IcicleCore};

\item
We present production benchmarks of our generated program, which compare favourably against standard unix utilities, and outperforms an existing R feature generation system by several orders of magnitude\sref{s:Benchmarks}.
\end{itemize}
